### 2021-06-17 manuscript/methodology.tex
Correcoes da Valeria pro relatorio finalizada.

Let $(x_i, y_i, z_i), i =  1,2,3,...N$, be a observed dataset in a region, the total-field anomaly can be approximated to:
\begin{equation}
\Delta T(x_i, y_i, z_i) = \hat{\mathbf{F}}^{\top} \mathbf{B}(x_i, y_i, z_i) \: ,
\label{eq:tfanomaly}
\end{equation}
where, $\mathbf{B}(x_i, y_i, z_i)$ is the crustal field, $\hat{\mathbf{F}}^{\top}$ is the transposed unitary vector with the main field directions, with $\hat{\mathbf{F}}$ described as:
\begin{equation}
\hat{\mathbf{F}} = \left[
\begin{array}{c}
F_x \\
F_y \\
F_z
\end{array} \right] = 
\left[
\begin{array}{c}
\cos(I_{0}) \, \cos(D_{0}) \\
\cos(I_{0}) \, \sin(D_{0}) \\
\sin(I_{0})
\end{array} \right],
\label{eq:unit_vector_F}
\end{equation}
where $I_{0}$ is the inclination and $D_{0}$ the declination of the main field , respectively.

Considering a uniform magnetized body with volume $v$ and a magnetization $\mathbf{m}$, the induced magnetic field is:
\begin{equation}
\mathbf{B}(x_i, y_i, z_i) = c_{m} \, \frac{\mu_{0}}{4\pi} \: \mathbf{M}(x_i, y_i, z_i) \: 
\mathbf{m} \: ,
\label{eq:induction-general}
\end{equation}
where, $\mu_{0} = 4\pi \, 10^{-7}$ H/m is the magnetic constant, 
$c_{m} = 10^{9}$ is a constant that transforms the induced magnetic field from Tesla (T) to nanotesla (nT) and 
$\mathbf{M}(x_i, y_i, z_i)$ is a matrix given by:
\begin{equation}
\mathbf{M}(x_i, y_i, z_i) =
\left[
\begin{array}{ccc}
\partial_{x_ix_i} \phi(x_i,y_i,z_i) & \partial_{xy} \phi(x_i,y_i,z_i) & 
\partial_{x_iz_i} \phi(x_i,y_i,z_i) \\
\partial_{x_iy_i} \phi(x_i,y_i,z_i) & \partial_{yy} \phi(x_i,y_i,z_i) & 
\partial_{y_iz_i} \phi(x_i,y_i,z_i) \\
\partial_{x_iz_i} \phi(x_i,y_i,z_i) & \partial_{yz} \phi(x_i,y_i,z_i) & 
\partial_{z_iz_i} \phi(x_i,y_i,z_i)
\end{array}
\right] \: ,
\label{eq:M}
\end{equation}
where, $\partial_{\alpha\beta} \phi(x_i,y_i,z_i,x_j,y_j,z_j)$ with $\alpha = x_i, y_i, z_i$ and 
$\beta = x_i, y_i, z_i$, are the second derivatives of the function $\phi(x_i,y_i,z_i,x_j,y_j,z_j)$
with respect to $x_i$, $y_i$ and $z_i$:

@@

\begin{equation}
\dfrac{1}{\sqrt{(x_i - x_j)^{2} + 
		(y_i - y_j)^{2} + (z_i - z_j)^{2}}} \, ,
\label{eq:r}
\end{equation}
and $x_j, y_j, z_j$ are the $j$th Cartesian coordinates of the volume elements where the integral (equation \ref{eq:phi}) is conducted within the body with volume $v$.

@@

It is possible to calculate the total-field anomaly $\Delta T(x_i, y_i, z_i)$ (equation \ref{eq:tfanomaly-general}) with the convolution between the harmonic function and the physical property:
\begin{equation}
\Delta T(x_{i}, y_{i}, z_{i})
= \int \limits_{-\infty}^{+\infty}
\int \limits_{-\infty}^{+\infty}
p(x_j, y_j, z_{c}) \,
\left[ c_{m} \, \frac{\mu_{0}}{4\pi} \,
\hat{\mathbf{F}}^{\top} \mathbf{H} \, 
\hat{\mathbf{h}} \right] \,
dx \, dy \: ,
\label{eq:tf-p-continuous-mag-positive}
\end{equation}
where $z_c$ is a constant representing the depth of the equivalent layer with $z_i < z_c$. The unitary vector $\hat{\mathbf{h}}$ is the magnetization directions of the equivalent sources across the layer:
\begin{equation}
\hat{\mathbf{h}} = \left[
\begin{array}{c}
{h}_x \\
{h}_y \\
{h}_z
\end{array} \right] = 
 \left[
\begin{array}{c}
\cos(I) \, \cos(D) \\
\cos(I) \, \sin(D) \\
\sin(I)
\end{array}
\right] \: ,
\label{eq:h_hat}
\end{equation}
and $\mathbf{H}$ is a $3 \times 3$ matrix that contains the second derivatives in relation to the observed dataset Cartesian coordinates $x_i, y_i, z_i$ as presented in equation \ref{eq:r}:
\begin{equation}
\mathbf{H} =
\left[
\begin{array}{ccc}
\partial_{xx} \frac{1}{r} & 
\partial_{xy} \frac{1}{r} & 
\partial_{xz} \frac{1}{r} \\
\partial_{xy} \frac{1}{r} & 
\partial_{yy} \frac{1}{r} & 
\partial_{yz} \frac{1}{r} \\
\partial_{xz} \frac{1}{r} & 
\partial_{yz} \frac{1}{r} & 
\partial_{zz} \frac{1}{r}
\end{array}
\right] \: =
\left[
\begin{array}{ccc}
H_{xx} & 
H_{xy} & 
H_{xz} \\
H_{xy} & 
H_{yy} & 
H_{yz} \\
H_{xz} & 
H_{yz} & 
H_{zz}
\end{array}
\right] \: ,
\label{eq:Hi}
\end{equation}
and the physical property $p(x_j, y_j, z_{c})$ represents the magnetic dipole moment distribution over the layer and has unit of $Am^2$.

Discretizing equation \ref{eq:tf-p-continuous-mag-positive}:
\begin{equation}
\Delta T(x_{i}, y_{i}, z_{i}) = \sum_{j=1}^{M} p_j a_{ij}\: ,
\label{eq:integral-sum_mag}
\end{equation}
where $j = 1,2,3,...,M$ are the equivalent sources discretizied and distributed across the layer and $a_{ij}$ is given by:
\begin{equation}
a_{ij}
= c_{m} \, \frac{\mu_{0}}{4\pi} \, \hat{\mathbf{F}}^{\top} \mathbf{H}_{ij} \: \hat{\mathbf{h}} \: .
\label{eq:aij_mag}
\end{equation}
The equation \ref{eq:aij_mag} of the magnetic equivalent layer can be written in the matrix form as:
\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the vector of total-field anomaly data, $\mathbf{A}$ is a matrix containing the elements given in equation \ref{eq:aij_mag}, also known as the sensitivity matrix and $\mathbf{p}$ is the vector of dipole moments of each equivalent source.
By using the normal equations, a conventional manner of solving linear systems, the estimative of the dipole moment parameters vector can be calculated from the total-field anomaly as:
\begin{equation}
\hat{\mathbf{p}} = \left( \mathbf{A}^{\top}\mathbf{A} %+ 
%\mu \, \mathbf{I} 
\right)^{-1}
\mathbf{A}^{\top} \mathbf{d}^{o} \: .
\label{eq:estimated-p-parameter-space}
\end{equation}
Equation \ref{eq:estimated-p-parameter-space} will be referenced throughout this work as the classical method for solving the equivalent layer.

@@

As an alternative from the clasical method of parameter estimative, the conjugate gradient (CG) is a well-known iterative method for solving symmetric and positive definite linear systems. By minimizing the quadratic form:
\begin{equation}
min \, \Phi(\mathbf{p}) = \frac{1}{2} \, \mathbf{p}^{\top} \, \mathbf{A} \, \mathbf{p} \,
- \mathbf{{d}^{o}}^{\top} \, \mathbf{p}\: ,
\label{eq:estimated-p-cg}
\end{equation}
it is possible to solve the system by constructing a basis of conjugate directions $c \in R^N$ \citep{aster2018parameter}. As the matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}) is not symmetric, instead we apply a general least square method minimizing:
\begin{equation}
min \, || \mathbf{A} \, \mathbf{p} - \mathbf{d}^o||_2 \: ,
\label{eq:estimated-p-cgls}
\end{equation}
by applying CG to the normal equations (equation \ref{eq:estimated-p-parameter-space}).

@@

	1 - If $it > 0$, let $\beta_{it} = \dfrac{{\mathbf{r}^{(it)}}^{\top} \, \mathbf{r}^{(it)}} {{\mathbf{r}^{(it - 1)}}^{\top} \, \mathbf{r}^{(it - 1)}}$

	2 - $\mathbf{c}^{(it)} = \mathbf{r}^{it} - \alpha_{it} \,\beta_{it} \, \mathbf{c}^{(it - 1)}$.

	3 - $\alpha_{it} = \dfrac{{||\mathbf{r}^{(it)}||^2_2}}{({\mathbf{c}^{(it)}}^{\top} \, \mathbf{A}^{\top})(\mathbf{A} \, \mathbf{c}^{(it)})}$.

	4 - $\hat{\mathbf{p}}^{(it + 1)} = \hat{\mathbf{p}}^{it} - \alpha_{it} \, \mathbf{c}^{(it)}$.

	5 - $\mathbf{s}^{(it + 1)} = \mathbf{s}^{it} - \alpha_{it} \, \mathbf{A} \, \mathbf{c}^{(it)}$.

@@

As shown in Algorithm 1, different from the classical method (equation \ref{eq:estimated-p-parameter-space}), no inverse matrix and no matrix-matrix product need to be calculate, but one matrix-vector multiplication is needed out of the loop and two matrix-vector multiplications, in steps 3 and 6, are necessary at each iteration. In this work, we will reduce the computational cost of the equivalent layer by substituting exactly these two matrix-vector products with a much more efficient algorithm.

@@

The notation used in this work will be the same as the one presented in \cite{takahashi2020convolutional}, where the authors described the structure of the symmetric Block-Toeplitz Toeplitz-Block matrix of the gravimetric equivalent layer by establishing a relation between the pair of \emph{matrix coordinates} $(x_i, y_i)$, $i = 1, ..., N$ or $(x_j, y_j)$, $j = 1, ..., M = N $ to a pair of \emph{grid coordinates} $(x_k, y_l)$ given as:
\begin{equation}
x_{i} \equiv x_{k} = x_{1} + \left[ k(i) - 1 \right] \, \Delta x \: , 
\label{eq:xi}
\end{equation}
and
\begin{equation}
y_{i} \equiv y_{l} = y_{1} + \left[ l(i) - 1 \right] \, \Delta y \: .
\label{eq:yi}
\end{equation}
In a $x$-\textit{oriented grid} the indices $i$ (or $j$) relate as integer functions of $k$ and $l$ by:
\begin{equation}
i(k, l) = (l - 1) \, N_{x} + k \quad ,
\label{eq:i-x-oriented}
\end{equation}
\begin{equation}
l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil
\label{eq:l-x-oriented}
\end{equation}
and
\begin{equation}
k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x} \quad .
\label{eq:k-x-oriented}
\end{equation}
For  $y$-\textit{oriented grid} they are given by:
\begin{equation}
i(k, l) = (k - 1) \, N_{y} + l \quad ,
\label{eq:i-y-oriented}
\end{equation}
\begin{equation}
k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \quad .
\label{eq:l-y-oriented}
\end{equation}
Figure \ref{fig:methodology} shows an example of a grid $N_{x} \times N_{y}$, where $N_{x} = 4$ and $N_{y} = 3$ demonstrating the relation between the \emph{matrix coordinates} with $k(i)$ and $l(i)$ depending on the orientation of the grid.

In a $x$-\textit{oriented grid} $q$ and $p$ gives the number of blocks ($Q = N_{y}$) and the number of elements of each block ($P = N_{x}$). They can be defined by the integer functions:
\begin{equation}
q(i, j) = \; l(i) - l(j)
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \; k(i) - k(j) \quad .
\label{eq:p-x-oriented}
\end{equation}
Equations \ref{eq:l-x-oriented} and \ref{eq:k-x-oriented} describes $l(i)$ and $l(j)$ and $k(i)$ and $k(j)$, respectively. When using $y$-oriented grids, $q$ and $p$ still defines the number of block and block elements, but $Q = N_{x}$ and $P = N_{y}$. But, the integer functions changes to:
\begin{equation}
q(i, j) = \; k(i) - k(j) 
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \; l(i) - l(j) \quad ,
\label{eq:p-y-oriented}
\end{equation}
where equation \ref{eq:k-y-oriented} now defines $k(i)$ and $k(j)$ and equation \ref{eq:l-y-oriented} defines $l(i)$ and $l(j)$. Note that equations \ref{eq:q-x-oriented}, \ref{eq:p-x-oriented}, \ref{eq:q-y-oriented} and \ref{eq:q-y-oriented} differs from the ones presented in \cite{takahashi2020convolutional} by the absence of the module.

In both $x$ or $y$-\textit{orientation}, matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}) can be rewritten by the indices $q$, $Q = -Q + 1,...,0,..., Q-1$ defining the number of its blocks:
\begin{equation}
\mathbf{A} = \begin{bmatrix}
\mathbf{A}^{0}   & \mathbf{A}^{-1} & \cdots          & \mathbf{A}^{-Q+1} \\
\mathbf{A}^{1}   & \mathbf{A}^{0}  & \mathbf{A}^{-1} & \vdots           \\ 
\vdots           & \mathbf{A}^{1}  & \ddots          & \mathbf{A}^{-1}   \\
\mathbf{A}^{Q-1} & \cdots          & \mathbf{A}^{1}  & \mathbf{A}^{0}                 
\end{bmatrix}_{N \times N} \: ,
\label{eq:BTTB_A}
\end{equation}
and by indice $p$, where each block has $P \times P$ elements $a^{q}_{p}$, $p = -P + 1,..., 0, \dots, P - 1$, :
\begin{equation}
\mathbf{A}_{q} = \begin{bmatrix}
a^{q}_{0}   & a^{q}_{-1} & \cdots     & a^{q}_{-P+1} \\
a^{q}_{1}   & a^{q}_{0}  & a^{q}_{-1} & \vdots           \\ 
\vdots      & a^{q}_{1}  & \ddots     & a^{q}_{-1}   \\
a^{q}_{P-1} & \cdots     & a^{q}_{1}  & a^{q}_{0}                 
\end{bmatrix}_{P \times P} \: ,
\label{eq:Aq_block}
\end{equation}
In general, matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}) is a non-symmetric BTTB, i.e., its blocks are non-symmetric ($\mathbf{A}^{-1} \neq \mathbf{A}^{1} $) and its elements also are non-symmetric ($a^{q}_{-1} \neq a^{q}_{1}$). Depending on specific values of the main field direction and the equivalent sources magnetization directions, matrix $\mathbf{A}$ can assume other structures, for example, when $\hat{\mathbf{F}} = [0, 0, 1]$ and $\hat{\mathbf{h}} = [0, 0, 1]$ it becomes symmetric. In this work we are considering the more commom situation for the matrix $\mathbf{A}$.

Also differently for the symmetric sensitivity matrix described by \cite{takahashi2020convolutional}, the non-symmetric BTTB matrix cannot be reconstructed only by its first column. Matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}) needs four: the first an last columns of the first column of blocks and the first and last columns of the last column of blocks. This has a physical implication in the equivalent layer which is not possible to use only one equivalent source to reprduce all elements of matrix $\mathbf{A}$, but instead takes four equivalent sources positioned at each corner. Figure \ref{fig:4_equivalent_sources} shows the positioning of the equivalent sources in a regular grid $N_x = 4$ $N_y = 3$ necessary to calculate the four columns capable of recover matrix $\mathbf{A}$.

@@

As pointed earlier in this work, the main improvement inside the CGLS method (Algorithm \ref{al:cgls-algorithm}) for estimating the parameter vector $\hat{\mathbf{p}}$ (equation \ref{eq:estimated-p-parameter-space}) is to substitute the matrix-vector multiplication $\mathbf{A}^{\top} \mathbf{s}^{(0)}$ out of the loop and primarily the two matrix-vector multiplications inside the loop at steps 3 an 6, $\mathbf{A} \, \mathbf{c}^{(it)}$ and $\mathbf{A}^{\top} \, \mathbf{s}^{(it + 1)}$, that is necessary at each iteration and takes most of its runtime.

>>>>

Let $(x_i, y_i, z_i), i =  1,2,3,...N$, be a observed dataset in a region considering a right-handed Cartesian coordinate system with the $x$-axis pointing north, $y$-axis pointing east and $z$-axis pointing downward. The total-field anomaly at the $i$th observation can be approximated to:
\begin{equation}
	\Delta T(x_i, y_i, z_i) = \hat{\mathbf{F}}^{\top} \mathbf{B}(x_i, y_i, z_i) \: ,
	\label{eq:tfanomaly}
\end{equation}
where, $\mathbf{B}(x_i, y_i, z_i)$ is the crustal field, $\hat{\mathbf{F}}^{\top}$ is the transposed unitary vector with the main field directions, with $\hat{\mathbf{F}}$ described as:
\begin{equation}
	\hat{\mathbf{F}} = \left[
	\begin{array}{c}
		F_x \\
		F_y \\
		F_z
	\end{array} \right] = 
	\left[
	\begin{array}{c}
		\cos(I_{0}) \, \cos(D_{0}) \\
		\cos(I_{0}) \, \sin(D_{0}) \\
		\sin(I_{0})
	\end{array} \right],
	\label{eq:unit_vector_F}
\end{equation}
where $I_{0}$ is the inclination and $D_{0}$ the declination of the main field, respectively.

Considering a uniform magnetized body with volume $v$ and a total magnetization vector $\mathbf{m}$, the induced magnetic field is:
\begin{equation}
	\mathbf{B}(x_i, y_i, z_i) = c_{m} \, \frac{\mu_{0}}{4\pi} \: \mathbf{M}(x_i, y_i, z_i) \: 
	\mathbf{m} \: ,
	\label{eq:induction-general}
\end{equation}
where, $\mu_{0} = 4\pi \, 10^{-7}$ H/m is the magnetic constant, 
$c_{m} = 10^{9}$ is a constant that transforms the induced magnetic field from Tesla (T) to nanotesla (nT) and 
$\mathbf{M}(x_i, y_i, z_i)$ is a $3 \times 3$ matrix given by:
\begin{equation}
	\mathbf{M}(x_i, y_i, z_i) =
	\left[
	\begin{array}{ccc}
		\partial_{xx} \phi(x_i,y_i,z_i) & \partial_{xy} \phi(x_i,y_i,z_i) & 
		\partial_{xz} \phi(x_i,y_i,z_i) \\
		\partial_{xy} \phi(x_i,y_i,z_i) & \partial_{yy} \phi(x_i,y_i,z_i) & 
		\partial_{yz} \phi(x_i,y_i,z_i) \\
		\partial_{xz} \phi(x_i,y_i,z_i) & \partial_{yz} \phi(x_i,y_i,z_i) & 
		\partial_{zz} \phi(x_i,y_i,z_i)
	\end{array}
	\right] \: ,
	\label{eq:M}
\end{equation}
where, $\partial_{\alpha\beta} \phi(x_i,y_i,z_i)$ with $\alpha = x, y, z$ and 
$\beta = x, y, z$, are the second derivatives of the function $\phi(x_i,y_i,z_i)$
with respect to $x$, $y$ and $z$:

@@

\begin{equation}
	\dfrac{1}{\sqrt{(x_i - x_j)^{2} + 
			(y_i - y_j)^{2} + (z_i - z_j)^{2}}} \,
	\label{eq:r}
\end{equation}
and $x_j, y_j, z_j$ are the $j$th Cartesian coordinates within the volume elements of the magnetized body with volume $v$, where the integral (equation \ref{eq:phi}) is conducted.

@@

Grounded on the equivalent layer principle it is possible to calculate the total-field anomaly $\Delta T(x_i, y_i, z_i)$ (equation \ref{eq:tfanomaly-general}) with the convolution between the harmonic function and the physical property:
\begin{equation}
\Delta T(x_{i}, y_{i}, z_{i})
= \int \limits_{-\infty}^{+\infty}
\int \limits_{-\infty}^{+\infty}
p(x_j, y_j, z_{c}) \,
\left[ c_{m} \, \frac{\mu_{0}}{4\pi} \,
\hat{\mathbf{F}}^{\top} \mathbf{H} \, 
\hat{\mathbf{h}} \right] \,
dx \, dy \: ,
\label{eq:tf-p-continuous-mag-positive}
\end{equation}
where $z_c$ is a constant representing the depth of the equivalent layer with $z_i < z_c$. The unitary vector $\hat{\mathbf{h}}$ is the magnetization directions of the equivalent sources over the layer:
\begin{equation}
\hat{\mathbf{h}} = \left[
\begin{array}{c}
{h}_x \\
{h}_y \\
{h}_z
\end{array} \right] = 
 \left[
\begin{array}{c}
\cos(I) \, \cos(D) \\
\cos(I) \, \sin(D) \\
\sin(I)
\end{array}
\right] \: ,
\label{eq:h_hat}
\end{equation}
where $I$ and $D$ are, respectively, the inclination and declination of the equivalent sources, $\mathbf{H}$ is a $3 \times 3$ matrix that contains the second derivatives in relation to the observed Cartesian coordinates $x, y, z$ as presented in equation \ref{eq:r}:
\begin{equation}
\mathbf{H} =
\left[
\begin{array}{ccc}
\partial_{xx} \frac{1}{r} & 
\partial_{xy} \frac{1}{r} & 
\partial_{xz} \frac{1}{r} \\
\partial_{xy} \frac{1}{r} & 
\partial_{yy} \frac{1}{r} & 
\partial_{yz} \frac{1}{r} \\
\partial_{xz} \frac{1}{r} & 
\partial_{yz} \frac{1}{r} & 
\partial_{zz} \frac{1}{r}
\end{array}
\right] \: =
\left[
\begin{array}{ccc}
H_{xx} & 
H_{xy} & 
H_{xz} \\
H_{xy} & 
H_{yy} & 
H_{yz} \\
H_{xz} & 
H_{yz} & 
H_{zz}
\end{array}
\right] \: ,
\label{eq:Hi}
\end{equation}
and the physical property $p(x_j, y_j, z_{c})$ represents the magnetic dipole moment distribution over the layer and has unit of $Am^2$.

Discretizing equation \ref{eq:tf-p-continuous-mag-positive} we get:
\begin{equation}
\Delta T(x_{i}, y_{i}, z_{i}) = \sum_{j=1}^{M} p_j a_{ij}\: ,
\label{eq:integral-sum_mag}
\end{equation}
where $j = 1,2,3,...,M$ are the equivalent sources discretized and distributed across the layer and $a_{ij}$ is given by:
\begin{equation}
a_{ij}
= c_{m} \, \frac{\mu_{0}}{4\pi} \, \hat{\mathbf{F}}^{\top} \mathbf{H}_{ij} \: \hat{\mathbf{h}} \: .
\label{eq:aij_mag}
\end{equation}
Equation \ref{eq:integral-sum_mag} can be written as:
\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the vector of total-field anomaly, $\mathbf{A}$ is a matrix containing the elements given by equation \ref{eq:aij_mag}, also known as the sensitivity matrix and $\mathbf{p}$ is the vector containing the dipole moments of each equivalent source. By using the normal equations:
\begin{equation}
	\mathbf{A}^{\top}\mathbf{A}{\mathbf{p}} = 
	\mathbf{A}^{\top} \mathbf{d}^{o} \:
	\label{eq:normal-equations}
\end{equation} 
which is a conventional manner of solving linear systems, the estimative of the dipole moment parameters vector can be calculated from the total-field anomaly as:
\begin{equation}
\hat{\mathbf{p}} = \left( \mathbf{A}^{\top}\mathbf{A} %+ 
%\mu \, \mathbf{I} 
\right)^{-1}
\mathbf{A}^{\top} \mathbf{d}^{o} \:
\label{eq:estimated-p-parameter-space}
\end{equation}
where $\mathbf{d}^{o}$ is the observed total-field anomaly. Equation \ref{eq:estimated-p-parameter-space} will be referenced throughout this work as the classical method for solving the equivalent layer.

@@

As an alternative from the clasical method of parameter estimative, the conjugate gradient (CG) is a well-known iterative method for solving linear systems with symmetric positive definite matrices. By minimizing the quadratic form:
\begin{equation}
\, \Phi(\mathbf{p}) = \frac{1}{2} \, \mathbf{p}^{\top} \, \mathbf{A} \, \mathbf{p} \,
- \mathbf{{d}^{o}}^{\top} \, \mathbf{p}\: ,
\label{eq:estimated-p-cg}
\end{equation}
it is possible to solve the system by constructing a basis of conjugate directions $c \in R^N$ \citep{aster2018parameter}. As we are solving a general least square problem and matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}) is not symmetric, instead we minimize:
\begin{equation}
\, || \mathbf{A} \, \mathbf{p} - \mathbf{d}^o||_2 \: ,
\label{eq:estimated-p-cgls}
\end{equation}
by applying the conjugate gradient to the normal equations (equation \ref{eq:normal-equations}).

@@

	1 - If $it > 0$, let $\beta_{(it)} = \dfrac{{\mathbf{r}^{(it)}}^{\top} \, \mathbf{r}^{(it)}} {{\mathbf{r}^{(it - 1)}}^{\top} \, \mathbf{r}^{(it - 1)}}$

	2 - $\mathbf{c}^{(it)} = \mathbf{r}^{(it)} - \alpha_{(it)} \,\beta_{(it)} \, \mathbf{c}^{(it - 1)}$.

	3 - $\alpha_{(it)} = \dfrac{{||\mathbf{r}^{(it)}||^2_2}}{({\mathbf{c}^{(it)}}^{\top} \, \mathbf{A}^{\top})(\mathbf{A} \, \mathbf{c}^{(it)})}$.

	4 - $\hat{\mathbf{p}}^{(it + 1)} = \hat{\mathbf{p}}^{(it)} - \alpha_{(it)} \, \mathbf{c}^{(it)}$.

	5 - $\mathbf{s}^{(it + 1)} = \mathbf{s}^{(it)} - \alpha_{(it)} \, \mathbf{A} \, \mathbf{c}^{(it)}$.

@@

Different from the classical least-square solution (equation \ref{eq:estimated-p-parameter-space}), the CGLS solution (Algorithm 1) requires neither inverse matrix nor matrix-matrix product.The CGLS method only requires: one matrix-vector multiplication out of the loop and two matrix-vector multiplications, in steps 3 and 6, at each iteration.

In this work, we will reduce the computational cost of the equivalent layer by substituting exactly these two matrix-vector products with a much more efficient algorithm.

@@

Let us consider that the observed total-field anomaly is located on an $N_x \times N_y$ regular grid of points spaced by $\Delta_x$ and $\Delta_y$ along the $x$- and $y$-directions, respectively.
The notation used in this work will be the same as the one presented in \cite{takahashi2020convolutional}, where the authors described the structure of the symmetric Block-Toeplitz Toeplitz-Block matrix of the gravimetric equivalent layer. Here, we also establish a relation between the pair of \emph{matrix coordinates} $(x_i, y_i)$, $i = 1, ..., N$ or $(x_j, y_j)$, $j = 1, ..., M = N $ and a pair of \emph{grid coordinates} $(x_k, y_l)$ given as:
\begin{equation}
x_{i} \equiv x_{k} = x_{1} + \left[ k(i) - 1 \right] \, \Delta x \: , 
\label{eq:xi}
\end{equation}
and
\begin{equation}
y_{i} \equiv y_{l} = y_{1} + \left[ l(i) - 1 \right] \, \Delta y \: .
\label{eq:yi}
\end{equation}
In a $x$-\textit{oriented grid} the indices $i$ (or $j$) relate as integer functions of $k$ and $l$ by:
\begin{equation}
i(k, l) = (l - 1) \, N_{x} + k \: ,
\label{eq:i-x-oriented}
\end{equation}
\begin{equation}
l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil
\label{eq:l-x-oriented}
\end{equation}
and
\begin{equation}
k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x} \: .
\label{eq:k-x-oriented}
\end{equation}
For  $y$-\textit{oriented grid} they are given by:
\begin{equation}
i(k, l) = (k - 1) \, N_{y} + l \: ,
\label{eq:i-y-oriented}
\end{equation}
\begin{equation}
k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: ,
\label{eq:l-y-oriented}
\end{equation}
where in equations \ref{eq:l-x-oriented} to \ref{eq:l-y-oriented}, $\lceil .\rceil$ is the celing function.
Figure \ref{fig:methodology} shows an example of a grid $N_{x} \times N_{y}$, where $N_{x} = 4$ and $N_{y} = 3$ demonstrating the relation between the \emph{matrix coordinates} with $k(i)$ and $l(i)$ depending on the orientation of the grid.

The $N \times M$ sensitivity matrix  $\mathbf{A}$ (equation \ref{eq:aij_mag}) can be represented as a grid of $Q \times Q$ blocks $\mathbf{A}^q$, $q = -Q+1,...,0,..., Q-1$. Each block  $\mathbf{A}^q$ has $P \times P$ elements $a^{q}_p$ where $p = -P+1^,...,0,...,P-1$

In a $x$-\textit{oriented grid} $q$ and $p$ give the number of blocks ($Q = N_{y}$) and the number of elements of each block ($P = N_{x}$). They can be defined by the integer functions:
\begin{equation}
q(i, j) = \; l(i) - l(j)
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \; k(i) - k(j) \: ,
\label{eq:p-x-oriented}
\end{equation}
where equations \ref{eq:l-x-oriented} and \ref{eq:k-x-oriented} describe $l(i)$ and $l(j)$ and $k(i)$ and $k(j)$, respectively. When using $y$-oriented grids, $q$ and $p$ still define the number of block and block elements, but $Q = N_{x}$ and $P = N_{y}$. Moreover, the integer functions changes to:
\begin{equation}
q(i, j) = \; k(i) - k(j) 
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \; l(i) - l(j) \: ,
\label{eq:p-y-oriented}
\end{equation}
where equation \ref{eq:k-y-oriented} now defines $k(i)$ and $k(j)$ and equation \ref{eq:l-y-oriented} defines $l(i)$ and $l(j)$. Note that equations \ref{eq:q-x-oriented}, \ref{eq:p-x-oriented}, \ref{eq:q-y-oriented} and \ref{eq:q-y-oriented} differs from the ones presented in \cite{takahashi2020convolutional} by the absence of the module.

In both $x$- or $y$-\textit{orientation}, matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}) can be rewritten by the indices $q = -Q + 1,...,0,..., Q-1$ defining the number of its blocks:
\begin{equation}
\mathbf{A} = \begin{bmatrix}
\mathbf{A}^{0}   & \mathbf{A}^{-1} & \cdots          & \mathbf{A}^{-Q+1} \\
\mathbf{A}^{1}   & \mathbf{A}^{0}  & \mathbf{A}^{-1} & \vdots           \\ 
\vdots           & \mathbf{A}^{1}  & \ddots          & \mathbf{A}^{-1}   \\
\mathbf{A}^{Q-1} & \cdots          & \mathbf{A}^{1}  & \mathbf{A}^{0}                 
\end{bmatrix}_{N \times N} \: ,
\label{eq:BTTB_A}
\end{equation}
and by indice $p$, where each block has $P \times P$ elements $a^{q}_{p}$, $p = -P + 1,..., 0, \dots, P - 1$:
\begin{equation}
\mathbf{A}^{q} = \begin{bmatrix}
a^{q}_{0}   & a^{q}_{-1} & \cdots     & a^{q}_{-P+1} \\
a^{q}_{1}   & a^{q}_{0}  & a^{q}_{-1} & \vdots           \\ 
\vdots      & a^{q}_{1}  & \ddots     & a^{q}_{-1}   \\
a^{q}_{P-1} & \cdots     & a^{q}_{1}  & a^{q}_{0}                 
\end{bmatrix}_{P \times P} \: ,
\label{eq:Aq_block}
\end{equation}
In general, matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}) is a non-symmetric BTTB, i.e., its blocks are non-symmetric ($\mathbf{A}^{-Q+1} \neq \mathbf{A}^{Q-1} $) and its elements also are non-symmetric ($a^{q}_{-1} \neq a^{q}_{1}$). Depending on specific values of the main field direction and the equivalent sources magnetization directions, matrix $\mathbf{A}$ can assume other structures, for example, when $\hat{\mathbf{F}} = [0, 0, 1]$ and $\hat{\mathbf{h}} = [0, 0, 1]$ it becomes symmetric. In this work, we are considering the more commom situation for the matrix $\mathbf{A}$.

Also differently for the symmetric sensitivity matrix described by \cite{takahashi2020convolutional}, the non-symmetric BTTB matrix cannot be reconstructed only by its first column. The construction of the matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}) needs four columns: the first and last columns of the first column of blocks and the first and last columns of the last column of blocks. This has a physical implication in the equivalent layer which is not possible to use only one equivalent source to reprduce all elements of matrix $\mathbf{A}$, such as in the gravity case as demonstrared by \cite{takahashi2020convolutional}. Rather, in the magnetic case it takes four equivalent sources positioned at each corner of the equivalent layer. Figure \ref{fig:4_equivalent_sources} shows the positioning of the equivalent sources in a regular grid $N_x = 4$ $N_y = 3$ necessary to calculate the four columns capable of recover the matrix $\mathbf{A}$.

@@

As pointed earlier in this work, the main improvement inside the CGLS method (Algorithm \ref{al:cgls-algorithm}) for estimating the parameter vector $\hat{\mathbf{p}}$ (equation \ref{eq:estimated-p-parameter-space}) is to substitute the matrix-vector multiplication $\mathbf{A}^{\top} \mathbf{s}^{(0)}$ out of the loop and the two matrix-vector multiplications inside the loop at steps 3 an 6, $\mathbf{A} \, \mathbf{c}^{(it)}$ and $\mathbf{A}^{\top} \, \mathbf{s}^{(it + 1)}$, that is necessary at each iteration and takes most of its runtime.
